---
title: "legislature_housing_affordablity_data"
author: "Jorge R. Soldevila Irizarry"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Source central setup file
source(here::here("./scripts/00_project_setup.R"))

# Remove scientific notation
options(scipen = 999,
        tigris_use_cache = TRUE
)
```

# This scripts downloads and calculates Housing Costs for Units with a Mortgage and without a Mortgage for Both Hispanics and Total Population in NYS using the function stored in the project setup file (00_project_setup.R).

## The data used for the analysis comes from table B25140 and B25140I, S20506, S207, and DP04 from the US Census Bureau.

### Housing Cost Burden Calculations

```{r Calculate housing cost burden and other data, message=FALSE, warning=FALSE}

assembly_affordability <- build_housing_affordability(
  geography = "state legislative district (lower chamber)",
  state = "NY",
  year = 2024,
  survey = "acs5"
)
```

### Export assembly housing cost burden

```{r Export legislature housing cost data, message=FALSE, warning=FALSE}

export_housing_affordability(
  data = assembly_affordability,
  geography = "state legislative district (lower chamber)",
  state = "NY",
  year = 2024
)

st_write(assembly_affordability,
         dsn = file.path(here::here("output"),
                         "legislature_housing_affordability.gpkg"),
         delete_dsn = TRUE)
```

## Table Export

```{r Export selected variables for Data brief, message=FALSE, warning=FALSE}

assembly_affordability_total_table <- 
  assembly_affordability %>% 
  select(
    NAME,
    DP05_0001E, # Total Population
    DP04_0001E, # Total Housing Units
    DP04_0002E, # Occupied Housing Units
    DP04_0002PE, # Percent Occupied Housing Units
    DP04_0090E, # Owner-Occupied Housing Units
    DP04_0091E, # Owner-Occupied Housing Units with a Mortgage
    DP04_0092E, # Owner-Occupied Housing Units without a Mortgage
    DP04_0047E, # Total Renter Occupied Housing Units
    DP04_0091PE, # Percent Owner-Occupied Housing Units with a Mortgage
    DP04_0092PE, # Percent Owner-Occupied Housing Units without a Mortgage
    DP04_0047PE, # Percent Total Renter Occupied Housing Units
    # Median Household Income - Missing
    S2506_C01_024E, # Owner Occupied Housing Units with Mortgage Median Household Income
    S2507_C01_019E, # Owner Occupied Housing Units without Mortgage Median Household Income
    # Median Household Income for Renter-Occupied Housing Units - Missing
    S2506_C01_040E, # Owner Occupied Housing Units with Mortgage Monthly Housing Costs in Median Dollars
    S2507_C01_032E, # Owner Occupied Housing Units without Mortgage Monthly Housing Costs in Median Dollars
    DP04_0134E, #Median Gross Rent ($)
    B25140_001E,  # Total Housing Units
    B25140_002E,  # Total Owned Housing Units with Mortgage
    mort_30pct,
    mort_50pct,
    B25140_006E,  # Total Owned Housing Units without Mortgage
    nomort_30pct,
    nomort_50pct,
    B25140_010E,  # Total Rented Housing Units
    rent_30pct,
    rent_50pct,
    DP04_0115E,   # SMOCAPI with mortgage > 35%
    DP04_0124E,   # SMOCAPI without mortgage > 35%
    DP04_0142PE, #GRAPI > 35%
    S2506_C01_009E, # Owner Occupied Housing Units with Mortgage Median Home Value
    S2507_C01_010E, # Owner Occupied Housing Units without Mortgage Median Home Value
         ) %>% 
  st_drop_geometry() %>%
  arrange(NAME) %>%
  column_to_rownames("NAME") %>%
  t() %>%
  as.data.frame()

assembly_affordability_hisp_table <- 
  assembly_affordability %>% 
  select(
    NAME,
    DP05_0090E, # Total Hispanic Population
    pct_hisp,   # Percent Hispanic Population
    B25140I_001E,  # Total Hispanic Housing Units
    B25140I_002E,  # Total Hispanic Owned Housing Units with Mortgage
    hisp_mort_30pct,
    hisp_mort_50pct,
    B25140I_006E,  # Total Hispanic Owned Housing Units without Mortgage
    hisp_nomort_30pct,
    hisp_nomort_50pct,
    B25140I_010E,  # Total Hispanic Rented Housing Units
    hisp_rent_30pct,
    hisp_rent_50pct) %>% 
  st_drop_geometry() %>%
  arrange(NAME) %>%
  column_to_rownames("NAME") %>%
  t() %>%
  as.data.frame()

write.csv(assembly_affordability_total_table,"assembly_affordability_total_table.csv")
write.csv(assembly_affordability_hisp_table,"assembly_affordability_hisp_table.csv")
```


