---
title: "senate_housing_affordablity_data"
author: "Jorge R. Soldevila Irizarry"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Source central setup file
source(here::here("./scripts/00_project_setup.R"))

# Remove scientific notation
options(scipen = 999,
        tigris_use_cache = TRUE
)
```

# This scripts downloads and calculates Housing Costs for Units with a Mortgage and without a Mortgage for Both Hispanics and Total Population in NYS using the function stored in the project setup file (00_project_setup.R).

## The data used for the analysis comes from table B25140 and B25140I, S20506, S207, and DP04 from the US Census Bureau.

### Housing Cost Burden Calculations

```{r Calculate housing cost burden and other data, message=FALSE, warning=FALSE}

senate_affordability <- build_housing_affordability(
  geography = "state legislative district (upper chamber)",
  state = "NY",
  year = 2024,
  survey = "acs5"
)
```

### Export County housing cost burden

```{r Export senate housing cost data, message=FALSE, warning=FALSE}
# 
# export_housing_affordability(
#   data = senate_affordability,
#   geography = "state legislative district (upper chamber)",
#   state = "NY",
#   year = 2024
# )

st_write(senate_affordability,
         dsn = file.path(here::here("output"),
                         "senate_housing_affordability.gpkg"),
         delete_dsn = TRUE)
```

