---
title: "county_housing_affordablity_data"
author: "Jorge R. Soldevila Irizarry"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Source central setup file
source(here::here("./scripts/00_project_setup.R"))

# Remove scientific notation
options(scipen = 999,
        tigris_use_cache = TRUE
)
```

# This scripts downloads and calculates Housing Costs for Units with a Mortgage and without a Mortgage for Both Hispanics and Total Population in NYS.

## The data used for the analysis comes from table B25140 and B25140I, S20506, S207, and DP04 from the US Census Bureau.

### Cost Burden

```{r Total Pop Housing Burden, message=FALSE, warning=FALSE}

# To obtain data from the U.S. Census we use tidycensus and the get_acs() function.

con_burden <- get_acs(
  geography = "county",
  state = "NY",
  year = 2024,
  survey = "acs5",
  table = "B25140",
  geometry = TRUE,
  output = "wide"
) %>%
  mutate(
    mort_30pct = round(B25140_003E/B25140_002E*100,2),
         mort_50pct = round(B25140_004E/B25140_002E*100,2),
         nomort_30pct = round(B25140_007E/B25140_006E*100,2),
         nomort_50pct = round(B25140_008E/B25140_006E*100,2),
         rent_30pct = round(B25140_011E/B25140_010E*100,2),
         rent_50pct = round(B25140_012E/B25140_010E*100,2)
    ) %>%
  select(GEOID,NAME,mort_30pct,mort_50pct,nomort_30pct,nomort_50pct,rent_30pct,
         rent_50pct)
```

### Display Burden Data

```{r 30% Household Burden Map, message=FALSE, warning=FALSE}

ggplot(
  con_burden) +
  geom_sf(aes(fill = mort_30pct)) +
  ggtitle(label = "Owned Units with a mortgage with Housing Costs Over 30% of Household Income") +
  theme_void() +
  scale_fill_viridis_c(option = "magma", 
                       direction = -1)

ggplot(
  con_burden) +
  geom_sf(aes(fill = mort_50pct)) +
  ggtitle(label = "Owned Units with a mortgage with Housing Costs Over 50% of Household Income") +
  theme_void() +
  scale_fill_viridis_c(option = "magma", direction = -1)
  
ggplot(
  con_burden) +
  geom_sf(aes(fill = nomort_30pct)) +
  ggtitle(label = "Owned Units without a mortgage with Housing Costs Over 30% of Household Income") +
  theme_void() +
  scale_fill_viridis_c(option = "magma", direction = -1)

ggplot(
  con_burden) +
  geom_sf(aes(fill = nomort_50pct)) +
  ggtitle(label = "Owned Units without a mortgage with Housing Costs Over 50% of Household Income") +
  theme_void() +
  scale_fill_viridis_c(option = "magma", direction = -1)

ggplot(
  con_burden) +
  geom_sf(aes(fill = rent_30pct)) +
  ggtitle(label = "Rented Units with Housing Costs Over 30% of Household Income") +
  theme_void() +
  scale_fill_viridis_c(option = "magma", direction = -1)

ggplot(
  con_burden) +
  geom_sf(aes(fill = mort_50pct)) +
  ggtitle(label = "Rented Units with Housing Costs Over 50% of Household Income") +
  theme_void() +
  scale_fill_viridis_c(option = "magma", direction = -1)
```

### Monthly Costs for Units with a Mortgage

```{r Cost with Mortgage, message=FALSE, warning=FALSE}

con_cost_wth_mort <- get_acs(
  geography = "county",
  state = "NY",
  year = 2024,
  survey = "acs5",
  table = "S2506",
  geometry = TRUE,
  output = "wide"
) %>%
  select(GEOID, 
         S2506_C01_001E, # Total Owner Occupied Housing Units with Mortgage
         S2506_C01_009E, # Owner Occupied Housing Units with Mortgage Median Home Value
         S2506_C01_024E, # Owner Occupied Housing Units with Mortgage Median Household Income
         S2506_C01_027E, # Home Value to Household Income Ratio 3 to 3.9
         S2506_C01_028E, # Home Value to Household Income Ratio 4 or >
         S2506_C01_040E, # Owner Occupied Housing Units with Mortgage Monthly Housing Costs in Median Dollars
         S2506_C02_066E  # Owner Occupied Housing Units with Mortgage Median Real Estate Taxes
         )
```

### Monthly Costs for Units without a Mortgage

```{r Cost without Mortgage, message=FALSE, warning=FALSE}

con_cost_wth_no_mort <- get_acs(
  geography = "county",
  state = "NY",
  year = 2024,
  survey = "acs5",
  table = "S2507",
  geometry = TRUE,
  output = "wide"
) %>%
  select(GEOID,
         S2507_C01_001E, # Total Owner Occupied Housing Units without Mortgage
         S2507_C01_010E, # Owner Occupied Housing Units without Mortgage Median Home Value
         S2507_C01_019E, # Owner Occupied Housing Units without Mortgage Median Household Income
         S2507_C01_022E, # Home Value to Household Income Ratio 3 to 3.9
         S2507_C01_023E, # Home Value to Household Income Ratio 4 or >
         S2507_C01_032E, # Owner Occupied Housing Units without Mortgage Monthly Housing Costs in Median Dollars
         S2507_C02_058E  # Owner Occupied Housing Units without Mortgage Median Real Estate Taxes
         )
```

### Housing Characteristics

```{r Cost with Mortgage, message=FALSE, warning=FALSE}

con_hous_charc <- get_acs(
  geography = "county",
  state = "NY",
  year = 2024,
  survey = "acs5",
  table = "DP04",
  geometry = TRUE,
  output = "wide"
) %>%
  select(GEOID, 
         DP04_0001E, # Total Housing Units
         DP04_0002E, # Occupied Housing Units
         DP04_0003E, # Vacant Housing Units
         DP04_0004E, # Homeowner Vacancy Rate
         DP04_0005E, # Renter Vacancy Rate
         DP04_0046E, # Total Owner Occupied Housing Units
         DP04_0047E, # Total Renter Occupied Housing Units
         DP04_0101E, # Median SMOC with Mortgage
         DP04_0109E, # Median SMOC without Mortgage
         DP04_0134E   # Median Gross Rent
  )
```

## Hispanic Population Burden

```{r Hispanic Pop Housing Burden, message=FALSE, warning=FALSE}

hisp_con_burden <- get_acs(
  geography = "county",
  state = "NY",
  year = 2024,
  survey = "acs5",
  table = "B25140I",
  geometry = TRUE,
  output = "wide"
) %>%
  mutate(
    hisp_mort_30pct = round(B25140I_003E/B25140I_002E*100,2),
         hisp_mort_50pct = round(B25140I_004E/B25140I_002E*100,2),
         hisp_nomort_30pct = round(B25140I_007E/B25140I_006E*100,2),
         hisp_nomort_50pct = round(B25140I_008E/B25140I_006E*100,2),
         hisp_rent_30pct = round(B25140I_011E/B25140I_010E*100,2),
         hisp_rent_50pct = round(B25140I_012E/B25140I_010E*100,2)
    ) %>%
  select(GEOID,hisp_mort_30pct,hisp_mort_50pct,hisp_nomort_30pct,
         hisp_nomort_50pct,hisp_rent_30pct,hisp_rent_50pct)
```

## Hispanic Majority and Plurality Calculations

```{r Hispanic Majority and Plurality Calculations, message=FALSE, warning=FALSE}

con_hisp_calc <- get_acs(
  geography = "county",
  state = "NY",
  year = 2024,
  survey = "acs5",
  table = "DP05",
  geometry = TRUE,
  output = "wide"
) %>%
  mutate(pct_hisp = DP05_0090E/DP05_0089E*100,
         pct_wht = DP05_0096E/DP05_0089E*100,
         pct_blk = DP05_0097E/DP05_0089E*100,
         pct_aian = DP05_0098E/DP05_0089E*100,
         pct_asi = DP05_0099E/DP05_0089E*100,
         pct_oth = DP05_0101E/DP05_0089E*100,
         hisp_maj = case_when(pct_hisp > 50 ~ "majority",
                              TRUE ~ "not moajority"),
         largest_other = pmax(pct_wht,
                              pct_blk,
                              pct_aian,
                              pct_asi,
                              pct_oth,
                              na.rm = TRUE),

         hisp_plurality = case_when(
         pct_hisp > largest_other ~ "largest group",
            TRUE ~ "not largest group"),
         largest_group = case_when(
           pct_hisp >= pmax(pct_wht, pct_blk, pct_aian, pct_asi, pct_oth, na.rm = TRUE) ~ "Hispanic",
           pct_wht  >= pmax(pct_blk, pct_aian, pct_asi, pct_oth, na.rm = TRUE) ~ "White",
           pct_blk  >= pmax(pct_aian, pct_asi, pct_oth, na.rm = TRUE) ~ "Black",
           pct_aian >= pmax(pct_asi, pct_oth, na.rm = TRUE) ~ "AIAN",
           pct_asi  >= pct_oth ~ "Asian",
           TRUE ~ "Other")
) %>%
  select(GEOID, DP05_0089E,
         pct_hisp,
         pct_wht,
         pct_blk,
         hisp_maj,
         hisp_plurality,
         largest_group)

```

## Join Data Frames and export to ESRi

```{r Join data and Export}

county_list <- list(con_burden,con_cost_wth_mort, con_cost_wth_no_mort,
                    con_hous_charc,hisp_con_burden,con_hisp_calc)

county_df <- county_list |>
  map(~ st_drop_geometry(.x)) |>
  reduce(left_join, by = "GEOID")

county_df <- con_burden |>
  select(GEOID, geometry) |>
  left_join(county_df, by = "GEOID") |>
  st_as_sf()
```

```{r Export joined data frame, message=FALSE, warning=FALSE}

st_write(
  county_df,
  here::here(
    "output",
    "county_housing_affordability.gpkg"
  ),
  delete_dsn = TRUE
)

```


