---
title: "county_housing_affordablity_data"
author: "Jorge R. Soldevila Irizarry"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Source central setup file
source(here::here("./scripts/00_project_setup.R"))

# Remove scientific notation
options(scipen = 999,
        tigris_use_cache = TRUE
)
```

# This scripts downloads and calculates Housing Costs for Units with a Mortgage and without a Mortgage for Both Hispanics and Total Population in NYS using the function stored in the project setup file (00_project_setup.R).

## The data used for the analysis comes from table B25140 and B25140I, S20506, S207, and DP04 from the US Census Bureau.

### Housing Cost Burden Calculations

```{r Calculate housing cost burden and other data, message=FALSE, warning=FALSE}

county_affordability <- build_housing_affordability(
  geography = "county",
  state = "NY",
  year = 2024,
  survey = "acs5"
)
```

### Export County housing cost burden

```{r Export county housing cost data, message=FALSE, warning=FALSE}

# export_housing_affordability(
#   data = county_affordability,
#   geography = "county",
#   state = "NY",
#   year = 2024
# )
```

## Display Burden Data

## Owned Units with a mortgage with Housing Costs Over 30% of Household Income

```{r Owner with Mortgage 30%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = county_affordability,
  variable = "mort_30pct",
  title = "Mortgage Cost Burden (>30%)"
)

```

## Owner with a Mortgage with Housing Costs Over 50% of Household Income

```{r Owner with Mortgage 50%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = county_affordability,
  variable = "mort_50pct",
  title = "Mortgage Cost Burden (>50%)"
)
```
