---
title: "senate_housing_affordablity_data"
author: "Jorge R. Soldevila Irizarry"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Source central setup file
source(here::here("./scripts/00_project_setup.R"))

# Remove scientific notation
options(scipen = 999,
        tigris_use_cache = TRUE
)
```

# This scripts downloads and calculates Housing Costs for Units with a Mortgage and without a Mortgage for Both Hispanics and Total Population in NYS using the function stored in the project setup file (00_project_setup.R).

## The data used for the analysis comes from table B25140 and B25140I, S20506, S207, and DP04 from the US Census Bureau.

### Housing Cost Burden Calculations

```{r Calculate housing cost burden and other data, message=FALSE, warning=FALSE}

senate_affordability <- build_housing_affordability(
  geography = "state legislative district (upper chamber)",
  state = "NY",
  year = 2024,
  survey = "acs5"
)
```

### Export County housing cost burden

```{r Export senate housing cost data, message=FALSE, warning=FALSE}
# 
# export_housing_affordability(
#   data = senate_affordability,
#   geography = "state legislative district (upper chamber)",
#   state = "NY",
#   year = 2024
# )

st_write(senate_affordability,
         dsn = file.path(here::here("output"),
                         "senate_housing_affordability.gpkg"),
         delete_dsn = TRUE)
```

## Display Burden Data Total Population

### Owned Units with a mortgage with Housing Costs Over 30% of Household Income

```{r Owner with Mortgage 30%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "mort_30pct",
  title = "Mortgage Cost Burden (>30%)"
)

```

### Owner with a Mortgage with Housing Costs Over 50% of Household Income

```{r Owner with Mortgage 50%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "mort_50pct",
  title = "Mortgage Cost Burden (>50%)"
)
```

### Owner without a Mortgage with Housing Costs Over 30% of Household Income

```{r Owner without Mortgage 30%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "nomort_30pct",
  title = "Mortgage Cost Burden (>30%)"
)
```

### Owner without a Mortgage with Housing Costs Over 50% of Household Income

```{r Owner without Mortgage 50%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "nomort_50pct",
  title = "Mortgage Cost Burden (>50%)"
)
```

### Rented Unit with Housing Costs Over 30% of Household Income

```{r Rented 30%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "rent_30pct",
  title = "Cost Burden (>30%)"
)
```

### Rented Unit with Housing Costs Over 50% of Household Income

```{r Rented 50%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "rent_50pct",
  title = "Cost Burden (>50%)"
)
```

### Total Households with Housing Costs Over 30% of Household Income

```{r Total Costs 30%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "tot_30pct",
  title = "Cost Burden (>30%)"
)
```

### Total Households with Housing Costs Over 50% of Household Income

```{r Total Costs 50%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "tot_50pct",
  title = "Cost Burden (>50%)"
)
```


## Display Burden Data Hispanic Populaiton

### Hispanic Owned Units with a mortgage with Housing Costs Over 30% of Household Income

```{r Hispanic Owner with Mortgage 30%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "hisp_mort_30pct",
  title = "Mortgage Cost Burden (>30%)"
)

```

### Hispanic Owner with a Mortgage with Housing Costs Over 50% of Household Income

```{r Hispanic Owner with Mortgage 50%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "hisp_mort_50pct",
  title = "Mortgage Cost Burden (>50%)"
)
```

### Hispanic Owner without a Mortgage with Housing Costs Over 30% of Household Income

```{r Hispanic Owner without Mortgage 30%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "hisp_nomort_30pct",
  title = "Mortgage Cost Burden (>30%)"
)
```

### Hispanic Owner without a Mortgage with Housing Costs Over 50% of Household Income

```{r Hispanic Owner without Mortgage 50%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "hisp_nomort_50pct",
  title = "Mortgage Cost Burden (>50%)"
)
```

### Hispanic Rented Unit with Housing Costs Over 30% of Household Income

```{r Hispanic Rented 30%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "hisp_rent_30pct",
  title = "Cost Burden (>30%)"
)
```

### Hispanic Rented Unit with Housing Costs Over 50% of Household Income

```{r Hispanic Rented 50%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "hisp_rent_50pct",
  title = "Cost Burden (>50%)"
)
```

### Hispanic Households with Housing Costs Over 30% of Household Income

```{r Hispanic Costs 30%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "hisp_tot_30pct",
  title = "Cost Burden (>30%)"
)
```

### Hispanic Households with Housing Costs Over 50% of Household Income

```{r Hispanic Costs 50%, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "hisp_tot_50pct",
  title = "Cost Burden (>50%)"
)
```


## Other Selected Housing Costs

###Units with a Mortgage

### Home Value to Household Income Ratio 3 to 3.9 Households with a Mortgage

```{r Ratio 3 to 3.9 with mortgage, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "ratio_less4.x",
  title = "Ratio 3 to 3.9"
)
```


### Home Value to Household Income Ration Over 4 Households with a Mortgage

```{r Ratio over 4 with mortgage, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "ratio_over4.x",
  title = "Ratio 3 to 3.9"
)
```


### Owner Occupied Housing Units with Mortgage Monthly Housing Costs in Median Dollars

```{r Median Housing costs with mortgage, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "S2506_C01_040E",
  title = "Median Dollars ($)"
)
```


### Owner Occupied Housing Units with Mortgage Median Real Estate Taxes

```{r Real Estate Taxes with mortgage, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "S2506_C02_066E",
  title = "Median Dollars ($)"
)
```


### Units without a mortgage

### Home Value to Household Income Ratio 3 to 3.9 Households without a Mortgage

```{r Ratio 3 to 3.9 no mortgage, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "ratio_less4.y",
  title = "Ratio 3 to 3.9"
)
```


### Home Value to Household Income Ration Over 4 Households without a Mortgage

```{r Ratio Over 4 No mortgage, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "ratio_over4.y",
  title = "Ratio 3 to 3.9"
)
```


### Owner Occupied Housing Units without Mortgage Monthly Housing Costs in Median Dollars

```{r Median Monthly costs no mortgage, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "S2507_C01_032E",
  title = "Median Dollars ($)"
)
```


### Owner Occupied Housing Units without Mortgage Median Real Estate Taxes

```{r Reatl estate taxes no mortgage, message=FALSE, warning=FALSE}

interactive_housing_map(
  data = senate_affordability,
  variable = "S2507_C02_058E",
  title = "Median Dollars ($)"
)
```


## Housing characteristics

### Median SMOC Households with a Mortgage

```{r Median Smoc with Mortgage, message=FALSE, warning=FALSE}
interactive_housing_map(
  data = senate_affordability,
  variable = "DP04_0101E",
  title = "Median Dollars ($)"
)
```

### Median SMOC Households without a Mortgage

```{r Median Smoc without Mortgage, message=FALSE, warning=FALSE}
interactive_housing_map(
  data = senate_affordability,
  variable = "DP04_0109E",
  title = "Median Dollars ($)"
)
```

### Median Gross Rent

```{r Median Gross Rent, message=FALSE, warning=FALSE}
interactive_housing_map(
  data = senate_affordability,
  variable = "DP04_0134E",
  title = "Median Dollars ($)"
)
```